Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-02-27 Publications/2015-NeuroImage/05-results.tex
paper version candidate

\subsection*{Proof of concept on digital phantoms}
\label{sec:results_phantom}
Using a simplified subset of the experimental instrumentation described in section
  \nameref{sec:experiments_evaluation}, we collected quantitative results on the digital
  phantoms.
To warp the simulated datasets, we randomly generated 150 different realizations of
  the distortion $U^{-1}_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
  $\times$ 2 resolutions) were run.
Parameter settings of every processing step in the experiments are made available along with
  the source code.
For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
  of the inner and the outer surfaces was computed.
The results showed a consistent and high accuracy, below voxel resolution.
\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
  to the two sets of resolutions of generated phantoms.
To support that the missregistration errors averaged per experiment was significantly
  under the resolution of the target image, we proceeded as follows.
First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
  normality.
All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
  non-normal with $p<0.001$.
Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
  correction for multiple comparisons ($N=150$).
Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
  for all the tests (4 phantom types $\times$ 2 resolutions).

@@

The distortions follow expression \eqref{eq:fieldmap} and hence, we easily reduced the
  error space to one dimension as follows.
The energy functional \eqref{eq:energy} was evaluated for the warpings given by
  a lineralized space of distortions
  $\hat{U}\{\vec{r}\} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
  with $\epsilon \in [-1.1, 1.1]$.
The result of this experiment is shown in \autoref{fig:energymap}.
The metric consistently displayed its minimum at $\epsilon=0.0$ (ground-truth location).

@@

\paragraph*{Evaluation \& cross-comparison}\label{sec:res_cc_evaluation}
%
Finally, we tested the performance of our registration method, and compared it to a competing
  method: the \emph{b0}-to-\gls*{t2} registration approach.
To this end, we integrated the registration software \emph{elastix} \citep{klein_elastix_2010},
  and reproduced the settings that internally uses \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
  \gls*{dti}.
A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
  was also implemented to run the correction experiment with the competing method.
First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
  the distorted \gls*{dmri} dataset.
Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
  reference image, and the \emph{b0} image as moving image, we performed the same correction
  experiment.

>>>>

\subsection*{Proof of concept on digital phantoms}
\label{sec:results_phantom}
Using a simplified subset of the experimental instrumentation described in section
  \nameref{sec:experiments_evaluation}, we collected quantitative results on the digital
  phantoms.
To warp the simulated datasets, we randomly generated 150 different realizations of
  the distortion $U_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
  $\times$ 2 resolutions) were run.
Parameter settings of every processing step in the experiments are made available along with
  the source code.
For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
  of the inner and the outer surfaces was computed.
The results showed a consistent and high accuracy, below voxel resolution.
\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
  to the two sets of resolutions of generated phantoms.
To support that the missregistration errors averaged per experiment was significantly
  under the resolution of the target image, we proceeded as follows.
First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
  normality.
All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
  non-normal with $p<0.001$.
Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
  correction for multiple comparisons ($N=150$).
Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
  for all the tests (4 phantom types $\times$ 2 resolutions).
Since statistical tests may not be conclusive enough, we also computed the confidence intervals
  (\autoref{tab:ci_phantom}).

\begin{table}
		\centering
		\footnotesize
    \begin{tabular}{llcc}
    Type      & Low-res.  & Hi-res. \\
    \hline
    ``gyrus'' & 0.58-0.60 & 0.18-0.36 \\
    ``L''     & 0.67-0.73 & 0.35-0.39 \\
    ``ball''  & 0.66-0.73 & 0.32-0.41 \\
    ``box''   & 0.68-0.71 & 0.35-0.39 \\
    \hline
    \end{tabular}
    \caption{All the 95\% confidence intervals (CI) were below the corresponding
      half-voxel resolution in all the phantom types.
    CIs reported in this table were computed using bootstrapping of $10^4$ samples,
      for each phantom type and resolution.
    Taking into account the non-normal distribution of errors, we used the median as location
  		statistic.}\label{tab:ci_phantom}
\end{table}

@@

The distortions follow expression \eqref{eq:fieldmap} and hence, we easily reduced the
  error space to one dimension as follows.
The energy functional \eqref{eq:energy} was evaluated for the warpings given by
  a lineralized space of distortions
  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
  with $\epsilon \in [-1.1, 1.1]$.
The results of this experiment are shown in \autoref{fig:energymap}.
The metric consistently displayed its minimum at $\epsilon=0.0$ (ground-truth location).

@@

\paragraph*{Evaluation \& cross-comparison}\label{sec:res_cc_evaluation}
%
Finally, we tested the performance of our registration method, and compared it to a competing
  method: the \emph{b0}-to-\gls*{t2} registration approach.
To this end, we integrated the registration software \emph{elastix} \citep{klein_elastix_2010},
  and reproduced the settings that internally uses \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
  \gls*{dti}.
A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
  was also implemented to run the correction experiment with the competing method.
First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
  the distorted \gls*{dmri} dataset.
Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
  reference image, and the \emph{b0} image as moving image, we registered both and
  projected the correct (undistorted) contours to the warped space using the resulting
  transform.
For both \emph{regseg} and the competing method, we computed the \gls*{swindex} \eqref{eq:swindex}
  of every surface.
Finally, to statistically compare the results from both methods we performed a one-way ANOVA test
  on the warping indices for three surfaces.
Following the nomenclature introduced in section \nameref{sec:human_connectome}, we selected
  $\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$ as surfaces of interest.
These interfaces are critical in connectivity analyses, and small errors in their location may
  lead to discarding fundamental paths between regions.
We also reported the 95\% CIs of the warping indices for those surfaces.
The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.

\begin{table}
		\centering
		\footnotesize
		\tabcolsep=0.05cm
    \begin{tabular}{L{1.8cm}ccc}
                            & Surf. 1  & Surf. 2 & Surf. 3 \\
    \hline
    CI \emph{regseg}        & 0.39 - 0.64 & 0.64 - 0.83 & 0.76 - 0.93 \\
    CI competing            & 2.36 - 3.37 & 2.65 - 3.19 & 2.82 - 3.40 \\
    \hline
    ANOVA (f-stat/p-value)  & (51.81/1.5e-06) & (148.25/8.1e-10) & (143.01/1.1e-09) \\
    \hline
    \end{tabular}
    \caption{All the 95\% CIs of the \gls*{swindex} distributions computed for the
      surfaces of interest were lower for \emph{regseg} with respect to the competing
      method.
    CIs reported in this table were computed using bootstrapping with mean as location
      statistic and $10^4$ samples.
    One-way-ANOVA tests indicated that there is a significant difference between our method and
      the competing method of choice.
    }\label{tab:results_real}
\end{table}

### 2015-02-27 Publications/2015-NeuroImage/06-discussion.tex
paper version candidate

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
In section \nameref{sec:methods_background}, we framed our method within the current
  state of art and defined the contributions in our approach.
We chose active contours without edges as we were interested in the homogeneous
  regions of diffusion properties within the brain.
Iteratively, the active contours are updated following the instant shape gradients
  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
The aproach is valid for image processing applications in which a precise segmentation is
  available (from an atlas or within subject anatomical data), and match any or all
  the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
To support our claim that images can be simultaneously segmented and registered to a
  reference coordinate system, we conducted evaluation experiments on digital
  simulated phantoms and real datasets.

The algorithm is designed for the proposed application on correcting distortions of
  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
  networks.
Connectome extraction protocols generally include the acquisition of \gls*{t1} images
  to provide with prior information about anatomy with high accuracy.
Our method solves in a single step a joint problem usually addressed in two steps.
The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data
  \citep{greve_accurate_2009}, whereas the second corrects for the nonlinear distortions
  derived from the inhomogenety of magnetic susceptibility across tissues
  \citep{jezzard_correction_1995}.

\paragraph*{Prospects}
Future extensions or improvements of \emph{regseg} similar to the evolution of existing
  alternative methodologies.
Incorporating knowledge about the distortion initializing our method with the theory-based
  correction using fieldmaps would be the most profitable.
The main pitfall using fieldmap correction is the challenging problem of unwrapping the
  range of the phase map, as it is clipped in the range $(-\pi, \pi]$.
The most extended solution to this is \emph{prelude} \citep{jenkinson_fast_2003}, and some
  alternatives have been proposed \citep{daga_susceptibility_2014} to circumvent the
  inherent unreliability of the method.
The \emph{regseg} algorithm would naturally correct missregistered areas due to errors in
  phase unwrapping, and errors derived to the important smoothing processes that are
  typically to clean up field maps.
This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach
  by \cite{irfanoglu_susceptibility_2011}, studied along with the virtues of using
  nonregular B-Spline grids.
This is also a potential extension to our deformation model.

>>>>

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
In section \nameref{sec:methods_background}, we framed our method within the current
  state of art and defined the contributions in our approach.
We chose active contours without edges as we were interested in the homogeneous
  regions of diffusion properties within the brain.
Iteratively, the active contours are updated following the instant shape gradients
  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
The aproach is valid for image processing applications in which a precise segmentation is
  available (from an atlas or within subject anatomical data), and match any or all
  the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that preclude voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
To support our claim that images can be simultaneously segmented and registered to a
  reference coordinate system, we conducted evaluation experiments on digital
  simulated phantoms and real datasets.

We designed \emph{regseg} as a method to correct \gls*{dmri} data for susceptibility
  distortions.
Connectome extraction protocols generally include the acquisition of \gls*{t1} images
  to provide with prior information about anatomy with high accuracy.
Our method solves in one single step a joint problem usually addressed in two steps:
  linear registration of \gls*{t1} and \gls*{dmri} to fuse the anatomical information on
  the diffusion data, and distortions correction using extra acquisitions.
Moreover, our method does not require any additional acquistition over the minimal protocol
  that only includes \gls*{t1} and \gls*{dmri} images, a very common scenario for
  historical datasets.

\paragraph*{Accuracy tests}
We proved that active contours without edges can be used to successfully drive a
  deformation supported by B-Spline functions in digital phantoms.
We randomly deformed 4 different phantom models mimicking three homogeneous regions,
  showing contrasts generated by a publicly available \gls*{mri} simulator.
After registration with \emph{regseg} we measured the Hausdorff distance between the
  projected contours using the ground-truth warping and our estimation.
We concluded that errors are significantly lower than voxel resolution.
We also assessed the 95\% intervals, yielding an aggregated interval of distances of
  $[0.58 - 0.73]$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
  $[0.18 - 0.41]$ in the high resolution phantoms ($1.0mm$ isotropic).
These aggregated intervals by resolution take the lowest of the lower boundary
  and the highest of the upper boundary of all CIs regardless the phantom type.
Therefore, we also conclude that the error is bounded and below the half of the
  voxel spacing to illustrate the accuracy of \emph{regseg}.

\paragraph*{Application on real data}
After a preliminary experiment to demonstrate the aptness of our segmentation model
  in section \nameref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
We processed 16 subjects downloaded from the \gls*{hcp} database with both \emph{regseg}
  and an in-house replication of a competing method.
This alternative approach solves distortion using \emph{b0}-to-\gls*{t2} registration.
Our method presented significantly different results (ANOVA tests are reported in
  \autoref{tab:results_real}), and the 95\% CIs revealed the high accuracy of
  \emph{regseg}.
The aggregated CI for \emph{regseg} was $[0.39 - 0.93] mm$, whereas the competing method
  yielded an aggregated CI of $[2.36 - 3.40] mm$.
Even though we carefully configured the competing method by using the same algorithm and
  extending the same settings from a widely used software, cross-comparison
  experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
Therefore, these results do not reflect that the \emph{regseg} software \emph{is} better than the
  competing method, but they are communicated to proof that \emph{regseg} is a solid option to
  correct \gls*{dmri} data for distortions, and any other potential application.

\paragraph*{Prospects}
Future extensions or improvements of \emph{regseg} similar to the evolution of existing
  alternative methodologies.
Incorporating knowledge about the distortion initializing our method with the theory-based
  correction using fieldmaps would be the most profitable.
The main pitfall using fieldmap correction is the challenging problem of unwrapping the
  range of the phase map, as it is clipped in the range $(-\pi, \pi]$.
The most extended solution to this is \emph{prelude} \citep{jenkinson_fast_2003}, and some
  alternatives have been proposed \citep{zhou_reliable_2009,daga_susceptibility_2014} to
  circumvent the inherent unreliability of the method.
The \emph{regseg} algorithm would naturally correct missregistered areas due to errors in
  phase unwrapping, and subestimations introduced by necessary smoothing processes generally
  applied to clean up field maps.
This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach
  by \cite{irfanoglu_susceptibility_2011}, studied along with the virtues of using
  nonregular B-Spline grids.
This is also a potential extension to our deformation model.

### 2015-02-27 Publications/2015-NeuroImage/supplemental.tex
paper version candidate

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-02-23 13:43:56

@@

\usepackage{aliascnt}
\usepackage{relsize}
\usepackage[procnames]{listings}

@@

% Listings style -----------------------------------------------------------------------------
\definecolor{pink}{RGB}{255,0,90}
\definecolor{comments}{RGB}{50,120,110}
\definecolor{string}{RGB}{160,0,0}
\definecolor{keywords}{RGB}{0,150,0}
\def\listingsfont{\ttfamily}
\def\listingsfontinline{\ttfamily}

\lstnewenvironment{bashcode}[1][]
  {\lstset{language=bash}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  %identifierstyle=\color{green},
  procnamekeys={def,class},
  morekeywords={models, lambda, forms, as, from},
  numbers=none,
  frame=single,
  xleftmargin= 10pt,
  xrightmargin= 10pt,
  framexleftmargin=10pt,
  frameround=tttt,
  fillcolor=\color{gray!10},
  backgroundcolor=\color{gray!10}
}}
{}

\lstnewenvironment{pythoncode}[1][]
  {\lstset{language=bash}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  %identifierstyle=\color{green},
  procnamekeys={def,class},
  morekeywords={models, lambda, forms, as, from},
  numbers=left,
  numberstyle=\smaller\color{black!60},
  stepnumber=1,
  numbersep=5pt,
  frame=single,
  xleftmargin= 40pt,
  xrightmargin= 20pt,
  framexleftmargin=20pt,
  frameround=tttt,
  fillcolor=\color{gray!10},
  backgroundcolor=\color{gray!10}
}}
{}

@@

It is possible to get the description of available options running
  \lstinline!regseg -h!:

@@

\paragraph{\emph{Nipype} interface}
Our registration algorithm is released with a \emph{nipype Interface} packaged in
  \lstinline!pyacwereg.interfaces.acwereg!.
This interface has been comprehensively used in the evaluation workflows.

@@

\subsection{Convergence evidencing}\label{sec:convergence_evidence}
In order to track the evolution of the registration process, several internal variables
  are saved in the JSON log-file.
Using the JSON log-file as input for the \emph{nipype Interface}
  \lstinline!pyacwereg.interfaces.ACWEReport!, it is straightforward to obtain
  a visual assessment document presenting the convergence.
Online checking is also possible as the algorithm writes to the standard output as well.

@@

@@

>>>>

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-02-27 15:21:03

@@

\usepackage{aliascnt}
\usepackage{relsize}
%\usepackage{inconsolata}
\usepackage[scaled]{beramono}
\usepackage[procnames]{listings}

@@

% Listings style -----------------------------------------------------------------------------
\definecolor{pink}{RGB}{255,0,90}
\definecolor{comments}{RGB}{50,120,110}
\definecolor{string}{RGB}{160,0,0}
\definecolor{keywords}{RGB}{0,150,0}
\definecolor{listingbg}{gray}{0.95}

\lstset{basicstyle=\ttfamily\smaller\relax}

\newcommand*{\codeinline}[1]{\colorbox{listingbg}{\lstinline!#1!}}
%\newcommand\codeinline{\lstinline}

\lstnewenvironment{bashcode}[1][]
  {\lstset{language={}}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  %identifierstyle=\color{green},
  morekeywords={regseg},
  numbers=none,
  frame=single,
  xleftmargin= 10pt,
  xrightmargin= 10pt,
  framexleftmargin=10pt,
  frameround=tttt,
  fillcolor=\color{listingbg},
  backgroundcolor=\color{listingbg}
}}
{}

\lstnewenvironment{pythoncode}[1][]
  {\lstset{language=bash}\lstset{%
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  breaklines=true,
  basicstyle=\ttfamily\smaller\relax,
  keywordstyle=\color{keywords}\bfseries,
  commentstyle=\color{comments}\itshape,
  stringstyle=\color{string},
  showstringspaces=false,
  identifierstyle=\color{black!70},
  procnamekeys={def,class},
  morekeywords={models, lambda, forms, as, from, import},
  numbers=left,
  numberstyle=\smaller\color{black!60},
  stepnumber=1,
  numbersep=5pt,
  frame=single,
  xleftmargin= 40pt,
  xrightmargin= 20pt,
  framexleftmargin=20pt,
  frameround=tttt,
  fillcolor=\color{gray!10},
  backgroundcolor=\color{gray!10}
}}
{}

@@

It is possible to get the description of available options running
  \codeinline{regseg -h}:

@@

\paragraph{\emph{Nipype} interface}
Our registration algorithm is released with a \emph{nipype Interface} packaged in
  \codeinline{pyacwereg.interfaces.acwereg}.
This interface has been comprehensively used in the evaluation workflows.

@@

\subsection{Convergence evidencing}\label{sec:convergence_evidence}
In order to track the evolution of the registration process, several internal variables
  are saved in the JSON log-file.
Using the JSON log-file as input for the \emph{nipype Interface}
  \codeinline{ACWEReport}, it is straightforward to obtain
  a visual assessment document presenting the convergence.
Online checking is also possible as the algorithm writes to the standard output as well.

@@

This work is supported by two \emph{nipype Workflows} in order to ensure the reproducibility
  of the results.
All the intermediate results and figures in this paper have been encapsulated into
  the workflows and are available in \url{http://dx.doi.org/zenodolink}.

An overview of the general workflow is presented in the Figure 2 of the paper.
However, this figure has been extremely simplified for the best of visualization.
In this section, we review the main elements of the evaluation pipelines.

\paragraph{Preprocessing}

\paragraph{Distortion}

@@

%{\color{red} {The main diverging points with respect to
%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
%but we replace the Gaussian filtering by other constraints
%studied in \citep{nagel_investigation_1986} to better the problem;
%3) optimization is applied in the spectral
%domain, observing anisotropic and inhomogeneous mappings along each direction.
%With respect \citep{guyader_combined_2011}, the main differences are
%the distance function, and the spectral solution to the optimization updates,
%as we shall cover in \autoref{sec:numerical_implementation}.}}

%In this paper we formulate the joint registration-segmentation
%problem as follows.
%such that the known contours in anatomical space $T$ optimally segment
%the diffusion space $D$.
%
%
%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
%we alternatively use \emph{shape-gradients}
%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.

%
%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
%  its associated feature vector in the following.
%
%In our application, these surfaces are precise tissue interfaces of interest extracted
%  from a high-resolution, anatomically correct reference volume using
%  the well-established
%
%In this paper we propose a novel registration framework to simultaneously
%solving the segmentation, distortion and cortical parcellation challenges,
%by exploiting as strong shape-prior the detailed morphology extracted
%from high-resolution and anatomically correct \gls{mri}.
%Indeed, hereafter
%we assume this segmentation problem in anatomical images is reliably and
%accurately solved with readily available tools (e.g.
%\citep{fischl_freesurfer_2012}).
%After global alignment with \gls{t1} using existing approaches, the remaining
%spatial mismatch between anatomical and diffusion space is due to susceptibility
%distortions.
%Finally, we need to establish precise spatial correspondence between the
%surfaces in both spaces, including the tangential direction for parcellation.
%Therefore, we can reduce the problem to finding the differences of spatial
%distortion in between anatomical and \gls{dwi} space.
%We thus reformulate the segmentation problem as an inverse problem, where we
%seek for an underlying deformation field (the distortion) mapping
%from the structural space into the diffusion space, such that the structural
%contours segment optimally the \gls{dwi} data.
%In the process, the one-to-one
%correspondence between the contours in both spaces is guaranteed, and projection
%of parcellisation to \gls{dwi} space is implicit and consistent.
%
%We test our proposed joint segmentation-registration model on two different
%synthetic examples.
%The first example is a scalar sulcus model, where the
%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
%segmented correctly thanks to the shape prior and its coupling with the inner,
%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
%The second case deals with more realistic \gls{dwi} data stemming from
%phantom simulations of a simplistic brain data.
%Again, we show that the
%proposed model successfully segments the \gls{dwi} data based on two derived
%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
%of the dense distortion field.
%
%The rest of this paper is organized as follows.
%First, in \autoref{sec:methods}
%we introduce our proposed model for joint multivariate segmentation-registration.
%Then we provide a more detailed description of the data and experimental setup in
%\autoref{sec:experiments}.
%We present results in \autoref{sec:results} and conclude
%in \autoref{sec:conclusion}.
%
%
%The properties of the reconstructed tensors and derived scalar maps have
%been studied by \cite{ennis_orthogonal_2006}. Based on their
%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
%considered complementary features, and therefore we selected them for the
%energy model \eqref{eq:complete_energy} in driving the
%registration-segmentation process.
%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
%the \gls{md} is more related to the \emph{magnitude} of the process:
%
%\begin{align}
%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
%\end{align}
%where $\lambda_i$ are the eigenvalues of the diffusion tensor
%associated with the diffusion signal $S(\vec{q})$. There exist
%two main reasons to justify their choice.
%First, they are well-understood and standardized in clinical routine.
%Second, together they contain most of the information that is
%usually extracted from the \gls{dwi}-derived scalar maps
%\cite{ennis_orthogonal_2006}.
%
